---
title: "Happiness correlation with open data"
output:
  html_document:
    fig_width: 12
    fig_height: 8
    theme: spacelab
    highlight: kate
---

```{r, message = F, warning = F}

library(dplyr)

# Read in data files from `open-data` and `world-happiness` datasets

open_data <- read.csv("C:/Users/KATHY/Desktop/countries.csv", stringsAsFactors=F)
happiness <- read.csv("C:/Users/KATHY/Desktop/2015.csv", stringsAsFactors=F)

# Rename from "Country Name" to just "Country" so it's easier to join
colnames(open_data)[2] <- "Country"
# Join the two dataset files on "Country"
open_data_happiness <- open_data %>%
  left_join(happiness, by = "Country") %>%
  mutate(Country = factor(Country)) %>%
  # Keep only columns I plan to use
  select(Country, Region, X2015.Score, Happiness.Score, Economy..GDP.per.Capita., 
         Family, Health..Life.Expectancy., Freedom, Trust..Government.Corruption., 
         Generosity, Dystopia.Residual)

# Give the columns nicer names now that our data is in one dataframe
colnames(open_data_happiness) <- c("Country", "Region", "Openness", "Happiness", "GDP", "Family", "Health", "Freedom", "Trust", "Generosity", "DystopiaResidual")

```

Now that I have the data roughly how I want it, let's have a quick peek. I really like this package called [`formattable`](https://github.com/renkun-ken/formattable) for presenting information in dataframes. Let's use it to look at the characteristics of the top 10 countries with the highest scores for their open data sharing policies:

```{r formattable_table}

library(formattable)
open_data_happiness %>%
  # Which countries are the most open?
  arrange(desc(Openness)) %>%
  # Round our numeric variables to two decimal places
  mutate_each(funs(round(., 2)), -c(Country, Region, Openness)) %>%
  head(10) %>%
  formattable(list(
    Openness = color_bar("yellow"),
    Happiness = color_bar("lightgreen"),
    GDP = color_bar("deepskyblue"),
    Family = color_bar("deepskyblue"),
    Health = color_bar("deepskyblue"),
    Freedom = color_bar("deepskyblue"),
    Trust = color_bar("deepskyblue"),
    Generosity = color_bar("deepskyblue"),
    DystopiaResidual = color_bar("deepskyblue")
  ), align = "l")

  

```

The top ten most "open" countries aren't localized to one or two regions but instead span the globe which is interesting. Do any of these countries surprise you? I was interested to see Colombia in fourth position; you can see the country scores lower relative to other top countries on some measures including "Trust" (Government Corruption). I investigated a bit and found [this report](http://opendatabarometer.org/3rdedition/regional-report/latin-america/) from the Open Data Barometer which does in fact call out Colombia as a regional leader in its 2016 analysis of open data readiness, implementation, and impact among Latin American countries.

I also suspect that the United States will get kicked down a few notches in coming years, sadly. Anyway, onto answering our main question:

### Are open countries happy countries?

Now we're ready to answer our question of whether countries that index highly for data openness are also home to happy people. Let's find out. The plot below will show the openness score and happiness score for each country as measured in 2015.

```{r open_data_happiness_plot, message = F, warning = F}

library(ggplot2)
library(ggthemes)
library(viridis)



ggplot(open_data_happiness, 
  aes(x = Openness, 
      y = Happiness)) +
  geom_point(aes(colour = Region),
             size = 2) +
  geom_smooth(method="lm") +
  labs(x = "Openness Score",
       y = "Happiness Score",
       title = "Are open data friendly countries happy countries?",
       subtitle = "Data openness and happiness by country in 2015") +
  scale_color_viridis(discrete = T) +
  theme_minimal() +
  theme(text = element_text(size=16))

  



```



**Overall, it looks like the answer is ... yes!** One thing that stands out from this plot is that Western European countries cluster above the linear fit line on "Happiness" while Sub-Saharan African countries are grouped together at the bottom left, falling lowest on the "Happiness" scale.



### What other measures are correlated with "Openness"?



Finally, what if we want to look at correlations between all of our variables? While we're exploring these datasets, let's spend some time creating a quick correlation matrix or "heatmap" for some additional insight.



```{r correlation_plot}

library(corrplot)

# Subset the data to include only the variables we want to look at

open_data_corr <- open_data_happiness %>%
  select(Openness, Happiness, GDP, Family, Health, 
         Freedom, Trust, Generosity, DystopiaResidual) %>%
  mutate(Openness = as.numeric(Openness))

od_corr <- cor(open_data_corr, use = "complete", method = "pearson")

corrplot(od_corr)

```
The one thing I want to highlight in this plot is the positive correlation between a country's happiness and its "Dystopia Residual". This seems confusing at a glance! From the dataset's description, however, we learn that the "Dystopia Residual" is a country's difference from "a hypothetical country that has values equal to the world's lowest national averages for each of the six factors". 




